---
title: "subsetting Lab"
author: "Alexej Gossmann"
date: "July 11, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

In this lab you can use the interactive console to explore but please record your commands here.  

Remember anything you type here can be "sent" to the console with: Cmd-Enter (OS-X) or Ctrl-Enter (Windows/Linux).

1. Check to see if you have the `mtcars` dataset (which is included in the `datasets` package)

```{r}
data(mtcars)
dim(mtcars)
head(mtcars)
#mtcars <- select(mtcars, name, everything())
```

2. What class is `mtcars`?

```{r}
class(mtcars)
```

3. How many observations (rows) and variables (columns) are in the `mtcars` dataset?

```{r}
paste(nrow(mtcars), "rows")
paste(ncol(mtcars), "columns")
```

4. Select the subset of rows that get more than 20 miles per gallon (mpg) of fuel efficiency - how many are there?

```{r}
morethan20mpg <- filter(mtcars, mpg > 20)
dim(morethan20mpg)
head(morethan20mpg)
```

```{r}
morethan20mpg <- mtcars[mtcars$mpg > 20, ]
dim(morethan20mpg)
head(morethan20mpg)
```

5. Select the subset of rows that get less than 16 miles per gallon (mpg) of fuel efficiency and have more than 100 horsepower (hp) - how many are there?

```{r}
mtcars.subset <- filter(mtcars, mpg < 16, hp > 100)
dim(mtcars.subset)
head(mtcars.subset)
```

```{r}
mtcars.subset <- mtcars[mtcars$mpg < 16 & mtcars$hp > 100, ]
dim(mtcars.subset)
head(mtcars.subset)
```

6. Create a subset of the data that only contains the columns:wt, qsec, and hp for only the cars that have 8 cylinder, and assign this object to `carsSub` - what are the dimensions of this dataset?

```{r}
carsSub <- mtcars %>% filter(cyl == 8) %>% select(wt, qsec, hp) 
dim(carsSub)
head(carsSub)
```

```{r}
carsSub <- mtcars[mtcars$cyl == 8, c("wt", "qsec", "hp")]
dim(carsSub)
head(carsSub)
```

7. Convert the column names of `carsSub` to all upper case

```{r}
colnames(carsSub) <- toupper(colnames(carsSub))
head(carsSub)
```

8. Re-order the rows of `carsSub` by weight in increasing order

```{r}
carsSub <- arrange(carsSub, desc(WT))
head(carsSub)
```

```{r}
carsSub <- carsSub[order(carsSub$WT, decreasing=TRUE), ]
head(carsSub)
```

9. Copy mtcars into a variable called `cars` and rename the `mpg` column in `cars` to `MPG`

```{r}
cars <- rename(mtcars, MPG = mpg)
dim(cars)
head(cars)
```

10. Subset the columns from mtcars that end in "p" and call it `pvars`

```{r}
cars <- select(cars, ends_with("p"))
head(cars)
```

